---
title: "MDS-subgroup gene expression"
date: "18/03/27"
output: pdf_document
author: "Elsa Bernard"
---

```{r, echo=FALSE, results="hide", include=FALSE}
library(ggplot2)
library(gplots)
source("../utils/ggstyles.R")
library(glmnet)
```

# Data

Here we describe the training and test sets
in order to make sure we do things properly!

Training data are data with response (=subgroups)
identified using CD34+ gene expressions,
and the predictive features here are BMMNC gene expressions.

Test data are data with only BMMNC expression.


```{r, echo=FALSE}
genes.paper = c("HOXB6","HOXB3","GPSM1","NYNRIN","SLMO1","ZNF662","SPINK2","MAPK12","HOXA3","AMN",
		"ROBO4","MYO5C","PRODH","UGCG","OLFM4","LCN2","CAST","WARS","ELL2",
		"CYSTM1","CYBB",
		"JHDM1D","OLR1","CAMP","CRISP3")
# -------------------------- #
# DATA
# -------------------------- #
bm = read.table("../../data/BMMNC_log2cpm.txt", sep="\t", stringsAsFactors=F)
colnames(bm) = as.vector(sapply(colnames(bm), function(x) strsplit(x,split="_BMMNC")[[1]][1]))
bm = bm[,grep("PV",colnames(bm))]
cd34 = read.table("../../data/CD34_log2cpm.txt", sep="\t", stringsAsFactors=F)
cd34 = cd34[,grep("PV",colnames(cd34))]
clinical = read.csv("../../data/Clinical_info.csv", stringsAsFactors=F)
clinical$ID = paste0("PV",clinical$ID)

training.samples = intersect(colnames(bm),colnames(cd34))
test.samples = colnames(bm)[!colnames(bm)%in%colnames(cd34)]

#print(paste("We have",length(training.samples),"training samples"))
#print(paste("We have",length(test.samples),"test samples"))

test.clinical = clinical[match(test.samples,clinical$ID),]
training.clinical = clinical[match(training.samples,clinical$ID),]

#table(training.clinical$WHO, training.clinical$subgroup)
#table(training.clinical$AML, training.clinical$subgroup)
#table(test.clinical$WHO, test.clinical$subgroup)
#table(test.clinical$AML, test.clinical$subgroup)

design = t(bm[genes.paper,])
test.design = design[test.samples,]
training.design = design[training.samples,]

training.response = clinical$subgroup[match(rownames(training.design),clinical$ID)]
```

```{r, echo=TRUE}
# number of samples in the training set:
nrow(training.design)
# number of samples in the test set:
nrow(test.design)
# number of subgroups (1 or 2) in the training set:
table(training.response)
```

# Paper findings

Below the logistic regression parameters explained in [Shiozawa et al., 2017].

```{r, echo=TRUE}
# selected genes for logistic regression:
genes.paper = c("HOXB6","HOXB3","GPSM1","NYNRIN","SLMO1","ZNF662","SPINK2","MAPK12","HOXA3","AMN",
		"ROBO4","MYO5C","PRODH","UGCG","OLFM4","LCN2","CAST","WARS","ELL2",
		"CYSTM1","CYBB",
		"JHDM1D","OLR1","CAMP","CRISP3")
# logistic regression parameters:
alpha.paper = 0.65
lambda.paper = 0.216
# we selected here only the above genes in the design:
ncol(training.design)
```

# Reproduce subgroup prediction

Before to tackle the gene-signature problem, 
we first want to reproduce the classification described in the paper.

We simply run a logistic regression with the described parameters.

We then look if the subgroup predictions on the test set are similar to ones that have been sent to me.


## Logistic with paper parameters

Attempt0:

```{r, echo=TRUE}
# train a unpenalized logistic regression with specified genes:
model.rep0 = glmnet(training.design, training.response,
		   family="binomial", lambda=0)
# the coefficient are different than the ones from the paper:
coef(model.rep0)
# test the fitted model:
class.rep0 = predict(model.rep0, newx = test.design, type="class")
# the predictions do not match the ones from the paper:
table(paste0("clinical_",test.clinical$subgroup), paste0("predicted_",class.rep0))
```

Attempt1:

```{r, echo=TRUE}
# train a penalized logistic regression with specified genes and parameters:
model.rep = glmnet(training.design, training.response,
		   family="binomial",
		   alpha=alpha.paper, lambda=lambda.paper)
# test the fitted model:
class.rep = predict(model.rep, newx = test.design, type="class")
# the predictions do not match the ones from the paper:
table(paste0("clinical_",test.clinical$subgroup), paste0("predicted_",class.rep))
```

## Logistic with tuned lambda

Attemp2:

```{r, echo=TRUE}
# train a penalized logistic regression with specified genes and alpha and tune lambda:
model.rep2 = cv.glmnet(training.design, training.response, family="binomial", alpha=alpha.paper)
# test the fitted model:
class.rep2 = predict(model.rep2, newx = test.design, type="class", s="lambda.min")
# the predictions do not match the ones from the paper:
table(paste0("clinical_",test.clinical$subgroup), paste0("predicted_",class.rep2))
```

# Conclusion

By simply running a logistic regression with the parameters from [Shiozawa et al., 2017],
I am not able to reproduce the subgroup classification of the test set.

I might be doing something stupid...

Would be nice if you could share the code as a baseline first!



```{r, echo=FALSE}
# -------------------------- #
# REPRODUCE
# -------------------------- #
#ddfit = as.data.frame( cbind(training.design, training.response) )
#ddfit$training.response = factor(ddfit$training.response)
#myformula = paste0("training.response ~ ", paste(colnames(training.design),collapse="+"))
#model.fit = glm(as.formula(myformula), data=ddfit, family=binomial, alpha=alpha.paper, lambda=lambda.paper)
# TRY 3
#toadd = c("Age","Blast","Hb","ANC","Plt","Cyto","Sex")
#model.matrix( ~ 0 + WHO, training.clinical)
#training.see = cbind( training.design , training.clinical[,toadd], model.matrix( ~ 0 + WHO, training.clinical))
#model.rep3 = cv.glmnet(as.matrix(training.see), training.response, family="binomial", alpha=alpha.paper)
#class.rep3 = predict(model.rep3, newx = as.matrix(cbind(test.design, test.clinical[,toadd], model.matrix( ~ 0 + WHO, test.clinical)[,-4])), type="class", s="lambda.min")
#table(test.clinical$subgroup, class.rep3)
```
